---
title: "mixture control dataset "
output: html_document
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

### mixture control dataset

To demonstrate the utility of the R package, we run the Snakemake pipeline on a [single-cell RNAseq dataset](https://github.com/LuyiTian/sc_mixology) which contains cells from 5 cancer cell lines mixed together: A549, H1975, H2228, H838 and HCC827 (Tian et al. 2019). We tested the combination of 6 different k.param (10,20,30,60,80,100), 7 different resolutions (0.001,0.005,0.01,0.05,0.1,0.2,0.3) and 5 different number of PCs (3,5,7,10,15) to retain after dimension reduction resulting in a total of 210 different parameter sets.  

To follow the analysis, you can download the data at [osf.io](https://osf.io/rfbcg/)

```{r}
library(scclusteval)
library(tidyverse)
library(patchwork)
library(Seurat)

# read in the seurat object
sc_10x_5cl_seurat<- readRDS("data/sc_mixology/sc_10x_5cl_seruat.rds")

subsample_idents<- readRDS("data/sc_mixology/gather_subsample.rds")

fullsample_idents<- readRDS("data/sc_mixology/gather_full_sample.rds")
```

### explore full dataset

```{r }
## how many PCs to include
ElbowPlot(sc_10x_5cl_seurat)

# a tibble with a list column
fullsample_idents

# what's the relationship of clusters between resolution 0.05, 0.1 and 0.3 with the same and k_param

fullsample_idents %>% mutate(id = row_number()) %>%
  filter(pc == 15, resolution == 0.05, k_param == 20)


fullsample_idents %>% mutate(id = row_number()) %>%
  filter(pc == 15, resolution == 0.1, k_param == 20)

fullsample_idents %>% mutate(id = row_number()) %>%
  filter(pc == 15, resolution == 0.3, k_param == 20)

## x-axis is resolution of 0.1, and y-axis is resolution of 0.05
PairWiseJaccardSetsHeatmap(fullsample_idents$original_ident_full[[55]],
                           fullsample_idents$original_ident_full[[60]],
                           show_row_dend = F, show_column_dend = F,
                           cluster_row = F, cluster_column =F)
```
cluster 4 split into cluster 4 and cluster 6
cluster 0 split into cluster 0 and 5


```{r}
## x-axis is resolution of 0.3, and y-axis is resolution of 0.05
PairWiseJaccardSetsHeatmap(fullsample_idents$original_ident_full[[55]],
                           fullsample_idents$original_ident_full[[70]],
                           show_row_dend = F, show_column_dend = F,
                           cluster_row = F, cluster_column =F)
```
cluster 4 split into cluster 4,7 
cluster 2 split into cluster 2,8
cluster 1 split into cluster 1,5

```{r}
## x-axis is resolution of 0.3, and y-axis is resolution of 0.1
PairWiseJaccardSetsHeatmap(fullsample_idents$original_ident_full[[60]],
                           fullsample_idents$original_ident_full[[70]],
                           show_row_dend = F, show_column_dend = F,
                           cluster_row = F, cluster_column =F)
```

We see as we increase the resolution, the number of clusters increase. 

Let's check how the clusters are splitting when we increase the resolution.
```{r}
## PC= 15, k.param = 20, resolution = 0.05
sc_10x_5cl_seurat<- 
  FindNeighbors(sc_10x_5cl_seurat, dims = 1:15, k.param = 20) %>%
  FindClusters(resolution = 0.05)

sc_10x_5cl_seurat<- 
  FindNeighbors(sc_10x_5cl_seurat, dims = 1:15, k.param = 20) %>%
  FindClusters(resolution = 0.1)

sc_10x_5cl_seurat<- 
  FindNeighbors(sc_10x_5cl_seurat, dims = 1:15, k.param = 20) %>%
  FindClusters(resolution = 0.3) 

p1<- DimPlot(sc_10x_5cl_seurat, reduction = "umap", label = TRUE, group.by = "RNA_snn_res.0.05") + ggtitle("res 0.05")

p2<- DimPlot(sc_10x_5cl_seurat, reduction = "umap", label = TRUE, group.by = "RNA_snn_res.0.1") + ggtitle("res 0.1")

p3<- DimPlot(sc_10x_5cl_seurat, reduction = "umap", label = TRUE, group.by = "RNA_snn_res.0.3") + ggtitle("res 0.3")

p4<- DimPlot(sc_10x_5cl_seurat, reduction = "umap", group.by = "cell_line", label = TRUE) 

p1 + p2 + p3 + p4
```


### Some other visualizations 

```{r}
## cluster number for each set of parameter
purrr::map_dbl(fullsample_idents$original_ident_full, function(x) length(unique(x)))

## how many cells identity change from one cluster to another
ClusterIdentityChordPlot(fullsample_idents$original_ident_full[[55]],
                           fullsample_idents$original_ident_full[[60]])
## cluster size plot
ClusterSizeBarplot(fullsample_idents$original_ident_full[[55]]) +
  theme_classic()

## change idents to cluster id when resolution is 0.1
Idents(sc_10x_5cl_seurat)<- sc_10x_5cl_seurat@meta.data$RNA_snn_res.0.1

CalculateSilhouette(sc_10x_5cl_seurat, dims = 1:15) %>% head()

silhouette_scores<- CalculateSilhouette(sc_10x_5cl_seurat, dims = 1:15)

sil_p1<- SilhouetteRainCloudPlot(silhouette_scores)

## check silhouette score with resolution 0.05
Idents(sc_10x_5cl_seurat)<- sc_10x_5cl_seurat@meta.data$RNA_snn_res.0.05
silhouette_scores<- CalculateSilhouette(sc_10x_5cl_seurat, dims = 1:15)
sil_p2<- SilhouetteRainCloudPlot(silhouette_scores)

sil_p1 + sil_p2
```

From the jaccard heatmap we know that:
cluster 4 split (res = 0.05) into cluster 4 and cluster 6 (res = 0.1)
cluster 0 split (res = 0.05) into cluster 0 and 5 (res = 0.1)

cluster 0 and cluster 5 have on average smaller silhouette score than the original cluster 0, suggesting that it is not a good idea to split A549 cells.

cluster 4 and 6 seems to have a higher silhouette score than the orignal cluster 4, suggesting that H1975 cells can be future split into two subclusters. From the `UMAP` plot, it does suggest cluster 4 and 6 are seperated.

## explore the subsampled data

```{r}
# a tibble with two list columns, note that the ident number is 80% of the full dataset
subsample_idents

## check for one of the subsample experiment, the cell identities 
## should be the same before and after clustering, but the cluster identities 
## should be different

subsample_idents %>% mutate(id = row_number()) %>%
  filter(pc == 15, resolution == 0.05, k_param == 20)

identical(names(subsample_idents$original_ident[[5401]]), names(subsample_idents$recluster_ident[[5401]]))

table(subsample_idents$original_ident[[5401]])
table(subsample_idents$recluster_ident[[5401]])

```

### Jaccard Raincloud plot for different resolutions

```{r}

subsample_idents_list<- subsample_idents %>% 
  group_by(pc, resolution, k_param) %>% 
  nest()

subsample_idents_list %>% ungroup() %>% mutate(id = row_number()) %>%
  filter(pc == 15, resolution == 0.05, k_param == 20)


subsample_idents_list$data[[55]]

## for the n times repeating, matching the clusters before and after reclustering
## and assign the jaccard for that cluster
AssignHighestJaccard(subsample_idents_list$data[[55]]$original_ident, 
                     subsample_idents_list$data[[55]]$recluster_ident)


JaccardRainCloudPlot(subsample_idents_list$data[[55]]$original_ident,
                          subsample_idents_list$data[[55]]$recluster_ident) + 
        geom_hline(yintercept = c(0.6, 0.75), linetype = 2) +
        xlab("cluster id w/ k=20 res=0.1 pc=15") 
        
```

with resolution of 0.05, Seurat finds 5 clusters which is optimal given that 5 cancer cell lines were mixed together. The jaccard indices are close to 1 for all clusters after subsampling and reclustering suggesting all 5 clusters are quite stable.


### make a Raincloud plot for every combination of the parameters.

```{r}

subsample_idents_list2<- subsample_idents_list %>%
  mutate(plot = map(data, ~JaccardRainCloudPlot(.x$original_ident, .x$recluster_ident) + geom_hline(yintercept = c(0.6, 0.75), linetype = 2)))


p1<- subsample_idents_list2 %>% 
  filter(resolution == 0.05, pc == 15, k_param ==20) %>%
  pull(plot) %>% `[[`(1)  + ggtitle("resolution = 0.05, pc = 15, k.param = 20")

p2<- subsample_idents_list2 %>% 
  filter(resolution == 0.1, pc == 15, k_param ==20) %>%
  pull(plot) %>% `[[`(1)  + ggtitle("resolution = 0.1, pc = 15, k.param = 20")


p1/ p2

## to save to disk, give a name for each pdf
## subsample_idents_list2<- mutate(subsample_idents_list2, file_name = 
##                                 paste0("PC_",pc, "_", ## "resolution_",resolution, "_", "k_", k_param, ".pdf"))

# save to disk
#walk2(subsample_idents_list2$file_name, subsample_idents_list2$plot, ggsave, width = 10, height = 6.5)
```

From the Jaccard raincloud plot, cluster 0,5 (res = 0.1) have dropped jaccard similarity index. cluster 4,6 (res = 0.1) also have dropped jaccard similarity index, suggesting that the orginal cluster 0 and cluster 4 (res - 0.05) should not be further splitted.

Increasing resolution will always give more clusters and whenever we observe a bimodal distribution of jaccard in a cluster when increasing the resolution, it indicates this cluster can be merged with a different cluster in the resampling and reclustering procedure.

### Assign stable clusters

As a rule of thumb, clusters with a mean/median stability score less than 0.6 should be considered unstable. scores between 0.6 and 0.75 indicate that the cluster is measuring a pattern in the data. clusters with stability score greater than 0.85 are highly stable (Zumel and Mount 2014). This task can be achieved using AssignStableCluster function in our R package. We observed for some datasets, the jaccard index follows  a bimodal distribution, so the mean or median may not be representative. As an alternative, we also calculate the percentage of subsampling with a jaccard greater than a cutoff (e.g. 0.85), which can be used to check stability assessments.


```{r}
## for one set of parameter: k=8, res=0.6, pc = 20
## return a list
AssignStableCluster(subsample_idents_list$data[[55]]$original_ident,
                    subsample_idents_list$data[[55]]$recluster_ident,
                    jaccard_cutoff = 0.8,
                    method = "jaccard_percent", 
                    percent_cutoff = 0.8)

# ?AssignStableCluster
## for all sets of parameters
stable_clusters<- subsample_idents_list %>%
  mutate(stable_cluster = map(data, ~ AssignStableCluster(.x$original_ident,
                                                          .x$recluster_ident,
                                                          jaccard_cutoff = 0.8,
                                                          method = "jaccard_percent", 
                                                          percent_cutoff = 0.8)))
```

### plot scatter plot for different parameters sets 

with y axis representing the number of stable clusters and total number of clusters.

```{r}
ParameterSetScatterPlot(stable_clusters = stable_clusters,
                        fullsample_idents = fullsample_idents,
                        x_var = "k_param",
                        y_var = "number",
                        facet_rows = "resolution",
                        facet_cols = "pc")


ParameterSetScatterPlot(stable_clusters = stable_clusters,
                        fullsample_idents = fullsample_idents,
                        x_var = "resolution",
                        y_var = "number",
                        facet_rows = "k_param",
                        facet_cols = "pc")

```

We see there are multiple parameter sets generate the optimal 5 clusters. This is a very artifical dataset. The plot looks more informative for real dataset.

### Calculate percentage of cells in stable clusters

```{r}
#?CalculatePercentCellInStable

stable_and_full<- left_join(stable_clusters, fullsample_idents)
CalculatePercentCellInStable(stable_and_full$original_ident_full[[1]],
                             stable_and_full$stable_cluster[[1]]$stable_cluster)

# as expected, all clusters are stable, and the percentage should be 100%
CalculatePercentCellInStable(stable_and_full$original_ident_full[[55]],
                             stable_and_full$stable_cluster[[55]]$stable_cluster)


stable_and_full<- stable_and_full %>% 
  mutate(precentage_in_stable = map2_dbl(original_ident_full, stable_cluster,
                                            function(x, y) CalculatePercentCellInStable(x,                                                                                      y$stable_cluster)))
```

### plot percentage cells in stable cluster

The `ParameterSetScatterPlot` function will calculate the percentage of cells in stable clusters
and plot a scatter/line plot.

```{r}
ParameterSetScatterPlot(stable_clusters = stable_clusters,
                        fullsample_idents = fullsample_idents,
                        x_var = "k_param",
                        y_var = "percentage",
                        facet_rows = "resolution",
                        facet_cols = "pc") +
  ggtitle("percentage of cells in stable clusters")
```

